Copulas and their potential for ecology

### 2019-10-01 Paper.Rmd
finished editing (for now) the Q1 sections, but Shyanolina needs to fix a couple things

We fit 16 bivariate copula families, exhibiting a variety of 
lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
of using a large collection of families was solely to include a 
variety of alternative dependence structures to have a robust model
selection and multi-model inference procedure. For that purpose, it 
is not important for the reader to
understand the details of these copulas,
and, additonally, these copulas have been 
described formally elsewhere [@Joe1997]; so 
we only briefly identify each family, say a few words about tail associations 
of the families, and provide pictorial descriptions in the Appendices.
We used several families that can exhibit positive lower-tail dependence 
(of strength depending on parameters) and 
zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
and survival BB6 familes. We used several families that can exhibit positive
upper-tail dependence (strength depends on parameters) and zero lower-tail
dependence: the survival Clayton, Gumbel, Joe and BB6 families. We used
families that show zero upper- and lower-tail dependence: the normal
and Frank families. These families have pdfs symmetric about the line
$v=-u+1$. We also used several families that can show both upper- and
lower-tail dependence, in relative amounts that depend on parameter
values: the BB1, survival BB1, BB7 and survival BB7 families. 
We also used the BB8 family, which shows zero lower-tail dependence,
and zero upper-tail dependence except for a boundary case for the parameters.
And we used the survival BB8 copula, which shows zero upper-tail dependence, 
and zero lower-tail dependence except for a boundary case.
"Survival" families are rotations of the copula with the similar name by
180 degrees. 
See @Joe1997 for details on all these families. We used the implementations
of these families provided in the `VineCopula` package for the R programming
language. See Figs \ref{figped_N}-\ref{figped_BB1} and \ref{SM-figped_SC}-\ref{SM-figped_SBB8} 
for visual depictions of the pdfs of these copulas and how
pdfs and tail dependence are influenced by the parameters.

@@

<!--***DAN: Notes on what needs to be done to this section, so far only getting up through the presentation of the bivariate results:
1) be more nuanced in the presentation, don't give the appearance of sweeping results that dont agree with the general conclusions under the rug
DONE 2) Add AIC and BIC columsn to table S3, and all analogous figures (e.g., S5, S7, S9, etc). Also reduce the number of digits shown after the decimal to 2 for AIC and BIC and 4 for the other columns.
3) Move Table S3 to an appropriate place in the main text. Choose the place based on where it is first cited in the main text.
DONE 4) Sort the rows of Table S2 (and all analogous tables) AIC, putting the lowest
AIC at the top. This should mean the highest AICw is at the top.
DONE 5a) move some of the content from appendices S6 and S7 into methods, 
5b) and move fig S21 and table S4 into main text. These are non-parametric results for
soil C and N. So we will be presenting the nonparam results more comprehensively in the main text.
6) That means you also need to summarize the nonparm results for the other 3 bivariate systems in Table 2. How to do this? Maybe put the cor results only, for Spearman only (or Kendall, not sure which to choose) into Table 2.  
7) Probably ought to combine tables S6, S8, S10 into one table, and fix it so
whenever we currently cite one of those, change it so we cite the combined table instead.
DONE 8) The nonparam stats are not estimators of tail dependence, they measure 
something else (though related). Make it clear we know this. It turns out there are estimators
of tail dependence. Do we want to mention that?
-->

@@

The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
C and N data
were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates
of the data,
separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
(Fig. \ref{fig_soilCN_nonparam}). Results showed that tail associations of data are stronger in 
both the lower and upper tails than would be expected under a null hypothesis of a normal copula (Fig. \ref{fig_soilCN_nonparam}B-F).
The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
on surrogates, separately in two 
comparisons (Table \ref{tab_soilCN_asym}). Results showed that
upper-tail association was significantly stronger than lower-tail association. These results
are consistent with the model selection results, but go beyond them by providing information
about statistical significance. Thus our results provide an
affirmative answer to Q1 for the soil C and N data; this answer is 
represented in Fig. \ref{MasterFigure} as a solid outline around "soil C and N"
in the right-most box in the middle row.
<!--***DAN: you may need to change the box around the Cedar Creek data-->

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes Q1 results 
for the bivariate datasets, with some details in the Appendices. The two 
variables were significantly related for all datasets 
(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
always emerged as the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, row 2). 
The normal copula was a poor fit compared to the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, rows 3-4), except for Cedar Creek, for which the AIC
difference between these fits was not meaningfully large (Table \ref{SM-tab_cc2000fit}).
Thus we answer the first part of Q1 in the affirmative for 3 of 4 of our bivariate datasets. 
Often, either lower- or upper-tail dependence statistics differed substantially from 0 
(Table \ref{tab_bivar_summary}, rows 7-8), and/or these statistics differed from each other (Table \ref{tab_bivar_summary}, row 9), 
helping to answering the second and third parts of Q1. Though again the model
selection results do not provide 
information on statistical significance of these differences, and are subject to the
caveat that, for some datasets, even the lowest-AIC copula was not an objectively good fit
(Table \ref{tab_bivar_summary}, rows 5 and 6), nonparametric results 
(Table \ref{tab_bivar_summary}, row 10) showed that tail associations
deviated significantly from what would be expected from a normal-copula
null model, and were also significantly asymmetric, except for the Cedar Creek data.
See Figs \ref{SM-fig_bmr_birds_nonparam}-\ref{SM-fig_cc2000_nonparam} for analogues to 
Fig. \ref{fig_soilCN_nonparam} for the bivariate datasets other than the soil C and N
data, and see Table \ref{SM-tab_3biv_npa_summary} for analogues to Table \ref{tab_soilCN_asym}.
The first three datasets had stronger upper- 
than lower-tail dependence, and the last dataset had the reverse, though non-significantly. Our 
generally affirmative (except for Cedar Creek), 
empirically based answer to Q1 for the bivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names of 
the soil C and N and bird and mammal body mass versus metabolic rate
datasets, in the middle row of boxes.
<!--***DAN: change the box for Cedar Creek-->
We did the same analysis 
with Cedar Creek data from other available years, 1996-1999 and 2007. For 1996 and 
2007, biomass and Shannon's index were independent; for 1997-1999, results 
were similar to 2000.

We present green spruce aphid abundance (Data section; Table \ref{tab_data_info}) results
as an example multivariate analysis.
Independence was rejected for each pair of the 10 sampling locations. 
Best-fitting (lowest-AIC) copulas were non-normal
for the large majority of location pairs (`r tab_multivar_summary$Green_spruce_aphid_count$num.non_N` of `r tab_multivar_summary$Green_spruce_aphid_count$num.pairs` pairs), 
and the AIC
of the normal copula minus the minimum AIC, averaged across 
location pairs, was `r tab_multivar_summary$Green_spruce_aphid_count$Avg.N_minus_bestfit.AIC`.
Whereas `r tab_multivar_summary$Green_spruce_aphid_count$Avg.N_minus_bestfit.AIC`
would be only a marginally meaningful AIC difference for a single
location pair, it is more meaningful as an average across many
pairs. Relatedly, and illustrating the concept here, 
the chances of getting `r tab_multivar_summary$Green_spruce_aphid_count$num.non_N` or more 
non-normal location pairs if the chances were equal of getting
a normal versus a non-normal result (and taking into account that
the location pair (A,B) will necessarily 
produce the same result as the pair
(B,A)) is low, 
$`r 1-pbinom(tab_multivar_summary$Green_spruce_aphid_count$num.non_N/2-1,45,0.5)`$. 
These results help answer the first part of Q1: non-normal
copula structure appears to be meaningfully common in
these data, even if not universal.
Goodness of fit tests in every case 
failed to reject the hypothesis that the AIC-best copula 
family was also 
an objectively adequate description of 
the data; i.e., the collection of 
copula families we used was sufficiently broad to characterize 
these data. Model-averaged lower- and upper-tail
dependence statistics had $2.5^{th}$ and $97.5^{th}$ quantiles
(`r tab_multivar_summary$Green_spruce_aphid_count$LT_0.025_CI`, `r tab_multivar_summary$Green_spruce_aphid_count$LT_0.975_CI`) and 
(`r tab_multivar_summary$Green_spruce_aphid_count$UT_0.025_CI`, `r tab_multivar_summary$Green_spruce_aphid_count$UT_0.975_CI`), respectively,
thereby commonly differing 
from what a normal copula would give (i.e., 0), and helping to
answer the second
part of Q1: these data have greater tail dependence 
(lower and upper) than would be expected from a normal-copula null hypothesis. We note however that our model selection 
procedures again do not reveal whether tail-dependence 
parameters are *significantly* different from $0$, and we refer 
the reader to nonparametric results below for information
on significance. 
Model-averaged lower- minus upper-tail dependence statistics 
were positive for all but a few location pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). The chances
of getting `r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
or more positive values, here, under a null hypothesis of equal chances 
for positive and negative values (and again accounting for the fact that location pairs (A,B) and (B,A) will show the same result) was
again low, 
$`r 1-pbinom(sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)/2-1,45,0.5)`$.
Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
than the spatial synchrony of outbreaks. This answers the third part of Q1.

@@

<!--***DAN: Shyamolina, can you please insert code to make what
is currently Table S7 appear here? Please also make it disappear from
the sup mat. I want to move it from there to here, please. I am
talking about the table with label tab_count_resamp. Please also
search both the main text and the supp mat for that label, and 
fix references - you may have to fix some prefixes (SM-, MT-), since
the table will have moved from sup mat to main text.-->
<!--Show the statistical results in a table for green spruce aphid abundance data-->

<!--multivariate summary-->
Table \ref{tab_multivar_summary} summarizes Q1 results for
multivariate data. Results supported the 
conclusions that non-normal copula structure, non-normal tail dependence, and asymmetric tail 
dependence were common, though not universal, answering 
Q1 in the affirmative for these data. 
Most location pairs were non-independent 
(Table \ref{tab_multivar_summary}, row 2). The large majority of non-independent location 
pairs had best-fitting copulas that were not the normal copula (Table \ref{tab_multivar_summary}, 
row 3), and AIC values of best-fitting copulas were, on average
across location pairs, between $2.714$ (*Ceratium furca* abundance data)
and $5.764$ (leaf-curling plum aphid first-flight data) lower than
AIC values for the normal copula (Table \ref{tab_multivar_summary}, 
row 4). Best-fitting copulas were nearly always considered an adequate fit 
(Table \ref{tab_multivar_summary}, row 5). Some of the datasets (green spruce aphid
abundance, \textit{Ceratium furca} abundance, methane-flux) showed stronger 
lower- than upper-tail dependence 
(Table \ref{tab_multivar_summary}, rows 6-8), whereas leaf-curling plum aphid first flight
data showed the reverse. All of the values in Table \ref{tab_multivar_summary}, rows 8 deviated highly significantly
from what would have been expected under a null hypothesis 
of equal chances for positive and negative values.

Asymmetry results were generally verified by nonparametric approaches,
with the exception of the methane-flux data. 
For instance, the $95\%$ confidence intervals of the
mean over pairs of sampling locations of the statistic $\cor_l-\cor_u$ were 
(`r unname(round(stat_aphid_ff$numericdf[10,c(3,4)],3))`)
for the leaf-curling plum aphid first flight data
(Table \ref{tab_multivar_summary}, row 9), 
indicating greater upper-tail dependence, and consistent with
the results of Table \ref{tab_multivar_summary}, row 8. 
For the *Ceratium furca* abundance data, confidence intervals were
(`r unname(round(stat_plankton_north_sea$numericdf[10,c(3,4)],3))`), 
indicating greater lower-tail dependence, and again 
consistent with the results of Table \ref{tab_multivar_summary} (row 9).
For the methane-flux data, confidence intervals were 
(`r unname(round(stat_methane$numericdf[10,c(3,4)],3))`);
the asymmetry in tail dependence revealed by the model selection
results for the methane-flux data was apparently not strong enough to
also be revealed by the nonparametric analyses. 
Analogues to Table \ref{tab_count_resamp} are in Tables
\ref{SM-tab_ff_resamp}-\ref{SM-tab_methane_resamp}.
Our generally
affirmative, empirically based answer to Q1 for the multivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names
of those datasets, in the middle row of boxes.
We also carried out the same analyses for abundance and first-flight data for the 18 
aphid species for which we had data other than the green spruce and leaf-curling plum aphids,
as well as for the 21 plankton taxa for which we had abundance data other than
\textit{Ceratium furca} (results not shown); results supported the conclusion that 
non-normal copula structure and tail dependence, and asymmetry of tail dependence are 
common.

@@

<!--***DAN: could remove the next para if you need to save words-->
<!--a para on stability of the total biomass time series in a community, and skewness - sets up Shya's later paper as future work-->
Our results showed that the skewness of the 
spatial average of local time series is influenced 
by their tail dependence. But the same logic should also apply to any collection of time
series, whether associated with locations in space or not. 
Another potential application is time series of all species from a single community, e.g., all
plants in a quadrat surveyed repeatedly over time. A huge 
literature has focused on synchrony versus 
compensatory dynamics between such time series, and 
the influence of interspecific relationships on the 
variability of community
properties such as total biomass. Typically, variability of community biomass 
is measured with the coefficient of variation, but skewness
may also be of interest because it can help characterize "spikiness" 
through time. Future work on copula 
structure of interspecific relationships in communities and its implications for community 
variability is likely to be valuable.

>>>>

We fit 16 bivariate copula families, exhibiting a variety of 
lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
of using a large collection of families was solely to include a 
variety of alternative dependence structures to have a robust model
selection and multi-model inference procedure. For that purpose, it 
is not important for the reader to
understand the details of these copulas,
and, additonally, these copulas have been 
described formally elsewhere [@Joe1997]; so 
we only briefly identify each family, say a few words about tail associations 
of the families, and provide pictorial descriptions in the Appendices.
We used several families that can exhibit positive lower-tail dependence 
(of strength depending on parameters) and 
zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
and survival BB6 familes. We used several families that can exhibit positive
upper-tail dependence (strength depends on parameters) and zero lower-tail
dependence: the survival Clayton, Gumbel, Joe and BB6 families. We used
families that show zero upper- and lower-tail dependence: the normal
and Frank families. These families have pdfs symmetric about the line
$v=-u+1$. We also used several families that can show both upper- and
lower-tail dependence, in relative amounts that depend on parameter
values: the BB1, survival BB1, BB7 and survival BB7 families. 
We also used the BB8 family, which shows zero lower-tail dependence,
and zero upper-tail dependence except for a boundary case for the parameters.
And we used the survival BB8 copula, which shows zero upper-tail dependence, 
and zero lower-tail dependence except for a boundary case.
"Survival" families are rotations of the copula with the similar name by
180 degrees. 
See @Joe1997 for details on all these families. We used the implementations
of these families provided in the `VineCopula` package for the R programming
language. See Figs \ref{figped_N}-\ref{figped_BB1} and \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
for visual depictions of the pdfs of these copulas and how
pdfs and tail dependence are influenced by the parameters.

@@

@@

The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
C and N data
were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates
of the data,
separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
(Fig. \ref{fig_soilCN_nonparam}). Results showed that tail associations of data are stronger in 
both the lower and upper tails than would be expected under a null hypothesis of a normal copula (Fig. \ref{fig_soilCN_nonparam}B-F).
The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
on surrogates, separately in two 
comparisons (Table \ref{tab_soilCN_asym}). Results showed that
upper-tail association was significantly stronger than lower-tail association. These results
are consistent with the model selection results, but go beyond them by providing information
about statistical significance. Thus our results provide an
affirmative answer to Q1 for the soil C and N data; this answer is 
represented in Fig. \ref{MasterFigure} as a solid outline around "soil C and N"
in the right-most box in the middle row.
<!--***DAN: you need to change the box around the Cedar Creek data-->

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes Q1 results 
for the bivariate datasets, with some details in the Appendices. The two 
variables were significantly related for all datasets 
(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
always emerged as the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, row 2). 
The normal copula was a poor fit compared to the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, rows 3-4), except for Cedar Creek, for which the AIC
difference between these fits was not meaningfully large.
Thus we answer the first part of Q1 in the affirmative for 3 of 4 of our bivariate datasets. 
Often, either lower- or upper-tail dependence statistics differed substantially from 0 
(Table \ref{tab_bivar_summary}, rows 7-8), and/or these statistics differed from each other (Table \ref{tab_bivar_summary}, row 9), 
helping to answering the second and third parts of Q1. Though again the model
selection results do not provide 
information on statistical significance of these differences, and are subject to the
caveat that, for some datasets, even the lowest-AIC copula was not an objectively good fit
(Table \ref{tab_bivar_summary}, rows 5 and 6), nonparametric results 
(Table \ref{tab_bivar_summary}, row 10) showed that tail associations
deviated significantly from what would be expected from a normal-copula
null model, and were also significantly asymmetric, except for the Cedar Creek data.
See Figs \ref{SM-fig_bmr_birds_nonparam}-\ref{SM-fig_cc2000_nonparam} for analogues to 
Fig. \ref{fig_soilCN_nonparam} for the bivariate datasets other than the soil C and N
data, see Tables \ref{SM-tab_bmr_birds_fit}-\ref{SM-tab_cc2000fit} for analogues to Table \ref{tab_soilCNfit},
and see Table \ref{SM-tab_3biv_npa_summary} for analogues to Table \ref{tab_soilCN_asym}.
The first three datasets had stronger upper- 
than lower-tail dependence, and the last dataset had the reverse, though non-significantly. Our 
generally affirmative (except for Cedar Creek), 
empirically based answer to Q1 for the bivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names of 
the soil C and N and bird and mammal body mass versus metabolic rate
datasets, in the middle row of boxes.
<!--***DAN: change the box for Cedar Creek-->
We did the same analysis 
with Cedar Creek data from other available years, 1996-1999 and 2007. For 1996 and 
2007, biomass and Shannon's index were independent; for 1997-1999, results 
were similar to 2000.

We present green spruce aphid abundance (Data section; Table \ref{tab_data_info}) results
as an example multivariate analysis.
Independence was rejected for each pair of the 10 sampling locations. 
Best-fitting (lowest-AIC) copulas were non-normal
for the large majority of location pairs (`r tab_multivar_summary$Green_spruce_aphid_count$num.non_N` of `r tab_multivar_summary$Green_spruce_aphid_count$num.pairs` pairs), 
and the AIC
of the normal copula minus the minimum AIC, averaged across 
location pairs, was `r tab_multivar_summary$Green_spruce_aphid_count$Avg.N_minus_bestfit.AIC`.
Whereas `r tab_multivar_summary$Green_spruce_aphid_count$Avg.N_minus_bestfit.AIC`
would be only a marginally meaningful AIC difference for a single
location pair, it is more meaningful as an average across many
pairs. Relatedly, and illustrating the concept here, 
the chances of getting `r tab_multivar_summary$Green_spruce_aphid_count$num.non_N` or more 
non-normal location pairs if the chances were equal of getting
a normal versus a non-normal result (and taking into account that
the location pair (A,B) will necessarily 
produce the same result as the pair
(B,A)) is low, 
$`r 1-pbinom(tab_multivar_summary$Green_spruce_aphid_count$num.non_N/2-1,45,0.5)`$. 
These results help answer the first part of Q1: non-normal
copula structure appears to be meaningfully common in
these data, even if not universal.
Goodness of fit tests in every case 
failed to reject the hypothesis that the AIC-best copula 
family was also 
an objectively adequate description of 
the data; i.e., the collection of 
copula families we used was sufficiently broad to characterize 
these data. Model-averaged lower- and upper-tail
dependence statistics had $2.5^{th}$ and $97.5^{th}$ quantiles (across
location pairs)
(`r tab_multivar_summary$Green_spruce_aphid_count$LT_0.025_CI`, `r tab_multivar_summary$Green_spruce_aphid_count$LT_0.975_CI`) and 
(`r tab_multivar_summary$Green_spruce_aphid_count$UT_0.025_CI`, `r tab_multivar_summary$Green_spruce_aphid_count$UT_0.975_CI`), respectively,
thereby commonly differing 
from what a normal copula would give (i.e., 0), and helping to
answer the second
part of Q1: these data have greater tail dependence 
(lower and upper) than would be expected from a normal-copula null hypothesis. We note however that our model selection 
procedures again do not reveal whether tail-dependence 
parameters are *significantly* different from $0$, and we refer 
the reader to nonparametric results below for information
on significance. 
Model-averaged lower- minus upper-tail dependence statistics 
were positive for all but a few location pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). The chances
of getting `r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
or more positive values, here, under a null hypothesis of equal chances 
for positive and negative values (and again accounting for the fact that location pairs (A,B) and (B,A) will show the same result) was
again low, 
$`r 1-pbinom(sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)/2-1,45,0.5)`$.
Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
than the spatial synchrony of outbreaks. This answers the third part of Q1.

@@

<!--multivariate summary-->
Table \ref{tab_multivar_summary} summarizes Q1 results for
multivariate data. Results supported the 
conclusions that non-normal copula structure, non-normal tail dependence, and asymmetric tail 
associations were common, though not universal, answering 
Q1 in the affirmative for these data. 
Most location pairs were non-independent 
(Table \ref{tab_multivar_summary}, row 2). The large majority of non-independent location 
pairs had best-fitting copulas that were not the normal copula (Table \ref{tab_multivar_summary}, 
row 3), and AIC values of best-fitting copulas were, on average
across location pairs, between $2.714$ (*Ceratium furca* abundance data)
and $5.764$ (leaf-curling plum aphid first-flight data) lower than
AIC values for the normal copula (Table \ref{tab_multivar_summary}, 
row 4). Best-fitting copulas were nearly always considered an adequate fit 
(Table \ref{tab_multivar_summary}, row 5). Some of the datasets (green spruce aphid
abundance, \textit{Ceratium furca} abundance, methane-flux) showed stronger 
lower- than upper-tail dependence 
(Table \ref{tab_multivar_summary}, rows 6-8), whereas leaf-curling plum aphid first flight
data showed the reverse. All of the values in Table \ref{tab_multivar_summary}, rows 8 deviated highly significantly
from what would have been expected under a null hypothesis 
of equal chances for positive and negative values.

Asymmetry results were generally verified by nonparametric approaches,
with the exception of the methane-flux data. 
For instance, the $95\%$ confidence intervals of the
mean over pairs of sampling locations of the statistic $\cor_l-\cor_u$ were 
(`r unname(round(stat_aphid_ff$numericdf[10,c(3,4)],3))`)
<!--***DAN: Shyamolina, these do not match with the value in table 6! Please check BOTH values and make them correct, and make them agree! You seemed to have used the values for the P statistic by mistake.-->
for the leaf-curling plum aphid first flight data
(Table \ref{tab_multivar_summary}, row 9), 
indicating greater upper-tail dependence, and consistent with
the results of Table \ref{tab_multivar_summary}, row 8. 
For the *Ceratium furca* abundance data, confidence intervals were
(`r unname(round(stat_plankton_north_sea$numericdf[10,c(3,4)],3))`),
<!--***DAN: Shyamolina, these do not match with the value in table 6! Please check BOTH values and make them correct, and make them agree! You seemed to have used the values for the P statistic by mistake.-->
indicating greater lower-tail dependence, and again 
consistent with the results of Table \ref{tab_multivar_summary} (row 9).
For the methane-flux data, confidence intervals were 
(`r unname(round(stat_methane$numericdf[10,c(3,4)],3))`);
<!--***DAN: Shyamolina, these do not match with the value in table 6! Please check BOTH values and make them correct, and make them agree! You seemed to have used the values for the P statistic by mistake.-->
the asymmetry in tail dependence revealed by the model selection
results for the methane-flux data was apparently not strong enough to
also be revealed by the nonparametric analyses. 
Analogues to Table \ref{tab_count_resamp} are in Tables
\ref{SM-tab_ff_resamp}-\ref{SM-tab_methane_resamp}.
Our generally
affirmative, empirically based answer to Q1 for the multivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names
of those datasets, in the middle row of boxes.
We also carried out the same analyses for abundance and first-flight data for the 18 
aphid species for which we had data other than the green spruce and leaf-curling plum aphids,
as well as for the 21 plankton taxa for which we had abundance data other than
\textit{Ceratium furca} (results not shown); results supported the conclusion that 
non-normal copula structure and tail dependence, and asymmetry of tail associations, are 
common.

@@

<!--a para on stability of the total biomass time series in a community, and skewness - sets up Shya's later paper as future work-->
Our results showed that the skewness of the 
spatial average of local time series is influenced 
by their tail dependence. But the same logic should also apply to any collection of time
series, whether associated with locations in space or not. 
Another potential application is time series of all species from a single community, e.g., all
plants in a quadrat surveyed repeatedly over time. A huge 
literature has focused on synchrony versus 
compensatory dynamics between such time series, and 
the influence of interspecific relationships on the 
variability of community
properties such as total biomass. Typically, variability of community biomass 
is measured with the coefficient of variation, but skewness
may also be of interest because it can help characterize "spikiness" 
through time. Future work on copula 
structure of interspecific relationships in communities and its implications for community 
variability is likely to be valuable.
